\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage[T1]{fontenc}
\author{Remy Jaspers - 4499336}
\title{CAO HW4}
\begin{document}
	\maketitle
	\section*{1a}
	The lower 16 bits are put into the sign extender, which in this instruction is: \\
	0000 1000 0010 0101. \\
	The sign bit is copied to fill the remaining 16 bits, thus the output of the sign extend unit will be:\\
	0000 0000 0000 0000 0000 1000 0010 0101. \\
	At the input of the Adder, this value will be shifted left two bits, thus the value will become:\\
	0000 0000 0000 0000 0010 0000 1001 0100\\\\
	The shift left 2 unit will take the lower 26 bits of the instruction and shift these left two bits. The input for this unit is:\\
	1 1000 0010 0000 1000 0010 0101 \\
	Shifted left twice to produce:\\
	1 1000 0010 0000 1000 0010 0101 00\\\\
	This will be the output of the shift left 2 unit. The Jump address will be calculated from concatenating PC+4 to this address. The current PC value is not given so we cannot calculate the upper four bits.\\\\

	\section*{1b}
	From figure 4.22, p. 269:
	\begin{table}[h]
		\centering
		\label{my-label}
		\begin{tabular}{|l|l|}
			\hline
		RegDst	&  1\\ \hline
		ALUSRc	& 0 \\ \hline
		MemtoReg & 0  \\ \hline
		RegWrite&  1 \\ \hline
		MemRead 	&  0\\ \hline
		MemWrite	&  0 \\ \hline
		Branch 	&  0\\  \hline
	 	ALUOp1	&  1\\ \hline
  		ALUOp2	&  0\\  \hline \end{tabular}
	\end{table}
	\section*{1c}
	As this is an R-format instruction, the new PC value will just be PC+4.
	\section*{1d}
	The instruction fetched from the instruction memory is: \\\\
	\verb|or $at, $t4, $v0| \\\\
	The values in these registers are provided. \$t4 = R12 = 16 and \$v0 = R2 = -128. These are the inputs to the main ALU.\\
	16 | -128 = -112 = 0xFFFFFF90 = 1111 1111 1111 1111 1111 111 1001 0000.\\\\
	The inputs for the Adder are (as given in 1a) 0x00002094, from the shift left 2 unit. The other input comes from the PC+4. Thus the adder performs the calculation 0x00002094 + (PC + 4).
	\section*{2a}
	\begin{table}[h]
		\centering
		\label{my-label}
		\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
			\hline
			Instruction & C1  & C2  & C3  & C4  & C5  &C6  &C7  & C8  & C9  & C10 & C11 & C12 & C13   \\ \hline
			LW R2, 0(R1) & IF  & ID  & EX  & MEM  & WB  &  &  &  &  &  &  &  &   \\ \hline
			NOP &  &  &  &  &  &  &  &  &  &  &  &  &    \\ \hline
			NOP &  &  &  &  &  &  &  &  &  &  &  &  &    \\ \hline
			AND R1, R2, R1 &  &  &  & IF& ID  & EX  & MEM & WB &  &  &  &  &    \\ \hline
			LW R3, 0(R2)&  &  &  &  &  IF & ID  & EX & MEM & WB &  &  &  &    \\ \hline
			LW R1, 0(R1)&  &  &  &  &  & IF & ID & EX & MEM  & WB  &  &  &  \\ \hline
			NOP &  &  &  &  &  &  &  &  &  &  &  &  &    \\ \hline
			NOP&  &  &  &  &  &  &  &  &  &  &  &  &    \\ \hline
			SW R1, 0(R2)&  &  &  &  &  &  &  &  & IF & ID & EX  & MEM  & WB    \\ \hline
		\end{tabular}
	\end{table}
\noindent	The AND instruction needs to wait for the write back stage to occur before it can compute the result using the ALU. Two NOPS are necessary here, because we assume no forwarding unit is available. Same for the store word instruction. We need to wait for the previous instructions write back to finish. Again two nops are inserted. The total number of cycles needed is 13.
\section*{2b}
It is not possible to reorder this code to reduce the number of stalls. There are too many dependencies.
\section*{2c}
Only 9, instead of 13 cycles will be needed because there won't be NOPS inserted, as there's no hazard detection unit. In this way however, data integrity cannot be guaranteed for the dependency on the R1 register in the AND instruction and the preceding LW instruction. As this is a load-use hazard, we can only forward this from MEM to EX but this is impossible to do in the same cycle.\\\\
The third instruction (LW R3, 0(R2)) will be correct as there's no dependency on registers used in the previous instruction. There is a hazard again between the fourth LW and SW instruction. Again, the data in R1 cannot be guaranteed to be correct when the SW instruction is executed, because we can only forward from MEM, as in the previous case. 
\section*{2d}
\begin{table}[h]
	\centering
	\label{my-label}
	\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
		\hline
		Instruction &C1&C2&C3&C4&C5&C6&C7&C8&C9&C10 \\ \hline
		LW R2, 0(R1) & IF  & ID  & EX  & MEM  & WB  &  &  &  &  &     \\ \hline
		AND=>NOP &  &  IF& ID & X  & X & X &  &  &  &          \\ \hline
		AND R1, R2, R1 &  &  & IF & ID& EX  & MEM  & WB&  &  &        \\ \hline
		LW R3, 0(R2)&  &  &  &  &  IF & ID  & EX & MEM & WB &         \\ \hline
		LW R1, 0(R1)&  &  &  &  &  & IF & ID & EX & MEM  & WB       \\ \hline	 \hline
			ForwardA&00 &00 &00&01 &00&  &&&&     \\ \hline
				ForwardB&  00& 00 &00  &00  & 00  &&&&&   \\ \hline
					Hazard Detection Unit&  0& 0 & 1  & 0 & 0   &&&&&  \\ \hline
	\end{tabular}
\end{table}
\noindent In cycle 3, the hazard detection unit finds that the ID/EX stage needs the output of MEM/WB from the previous instruction. In that cycle, ID/EX.registerRS == MEM/WB.registerRT and it enables the multiplexer for forwarding from MEM/WB to ID/EX so that the value will be available.The AND then becomes a NOP allowing for R2 of the LW to settle before use in the EX stage of the AND in cycle 5.
\\\\
\noindent 10 cycles are necessary to complete this code, as one NOP is needed to forward data from MEM of LW to the EX stage of the AND instruction.
\end{document} 